---
title: "Homework 2"
author: "Yuki Joyama"
date: "2023-10-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
$p$, the probability of having at least one dental checkup during a two-year period, is 0.73. 

(a) Let X be the the number of people (out of 56 random individuals) that have at least one dental checkup. X~Bin(56, 0.73)

```{r echo = F}
# calculation
a = choose(56, 40)*(0.73^40)*(0.27^16)
```

$P(X=40)=f(x)={n \choose x}p^x(1-p)^{n-x}={56 \choose 40}(0.73)^{40}(1-0.73)^{56-40}=$ `r round(a, 2)`  

Therefore, the probability that exactly 40 of these individuals will have at least one dental check up is **`r round(a, 2)`**

(b)
The probability that at least 40 of these individuals will have at least one dental checkup can be denoted as: 
$P(X\geq40) = 1 - P(X<40) = 1 - P(X\leq39)$

```{r echo = F, results = F}
# calculation
x = c(0:39)
p = 0.73
q = 1 - 0.73
c = 0

for (i in x){
  c = ifelse(i %in% 0:39, c + choose(56, i)*(p^i)*(q^(56-i)), c)
}

b = 1 - c 

# check using binomial distribution
pbinom(39, 56, 0.73)
```

$P(X\leq39)=P(X=0)+P(X=1)+...+P(X=39)=$ `r round(c, 2)`  
Therefore, $P(X\geq40)=1 - P(X\leq39)=$ **`r round(b, 2)`**


(c)
```{r echo = F}
# calculation
npq = 56*p*q
np = 56*p
nq = 56*q

# using normal distribution
a_c = dnorm(39.5, mean = np, sd = sqrt(npq)) 
b_c = 1 - pnorm(39.5, mean = np, sd = sqrt(npq))
```
$np=56 \times 0.73=$ `r np` > 10 and $nq=56 \times 0.73=$ `r nq` > 10  
Thus we can approximate X to the normal distribution X~N(`r round(np, 2)`, `r round(sqrt(npq), 2)`) where $P(X = 40)$ becomes $P(X = 40 - \frac{1}{2})$ and $P(X \geq 40)$ becomes $P(X \geq 40 - \frac{1}{2})$  

When X~N(`r round(np, 2)`, `r round(sqrt(npq), 2)`),  
$P(X = 39.5)=$ **`r round(a_c, 2)`**   
$P(X \geq 39.5)=$ **`r round(b_c, 2)`**

These are similar to the results of (a) and (b).

(d)
The expected value of X~Bin(56, 0.73) is:  
$\mu=E(X)=np=$ **`r round(np)`**

(e)  
The variance of X~Bin(56, 0.73) is:  
$\sigma^2=var(X)=np(1-p)=$ `r round(npq, 2)`  
Therefore, the standard deviation is **`r round(sqrt(npq), 2)`**

# Problem 2
Suppose the number of tornadoes in the U.S. follows a Poisson distribution with parameter $\lambda=6$ tornadoes per year.  
Let X denote the number of tornadoes in the U.S. per year. X~Poi(6)

(a)
```{r echo = F}
# calculation
ax0 = ((6^0)*(exp(1)^-6))/(1)  
ax1 = ((6^1)*(exp(1)^-6))/(1)  
ax2 = ((6^2)*(exp(1)^-6))/(2*1)  
```
The probability of having fewer than 3 tornadoes in the U.S. next year is   
$P(X<3)=P(X=0)+P(X=1)+P(X=2)$  
$P(X=x)=f(x)=\frac{\lambda^xe^{-\lambda}}{x!}$ so, 

$P(X=0)=\frac{6^0e^{-6}}{0!}=$ `r round(ax0, 2)`  
$P(X=1)=\frac{6^1e^{-6}}{1!}=$ `r round(ax1, 2)`  
$P(X=2)=\frac{6^2e^{-6}}{2!}=$ `r round(ax2, 2)`    

Therefore, $P(X<3)=$ **`r round(ax0 + ax1 + ax2, 2)`**

(b) 
```{r echo = F, results = F}
# calculation
b2 = ((6^3)*(exp(1)^-6))/(3*2*1)

# check
dpois(3, 6)
```
The probability of having exactly 3 tornadoes in the U.S. next year is  
$P(X=3)=\frac{6^3e^{-6}}{3!}=$ **`r round(b2, 2)`**

(c)
```{r echo = F, results = F}
# check
ppois(3, 6, lower.tail = F)
```
The probability of having more than 3 tornadoes in the U.S. next year is   
$P(X>3)=1-P(X\leq3)= 1-(P(X<3)+P(X=3))=$ **`r round(1 - (ax0+ax1+ax2+b2), 2)`**

# Problem 3
Assume the systolic blood pressure (SBP) of 20-29 year old American males is normally distributed with population mean 128.0 and population standard deviation 10.2.  

Let X denote the systolic blood pressure of 20-29 year old American males.   
Now,  
X~N(128.0, 10.2), $\mu=128.0$, $\sigma=10.2$  

(a)
The probability that a randomly selected American male between 20 and 29 years old has a SBP above 137.0 is $P(X>137.0)=1-P(X\leq137.0)$ 

$P(X<x)=\int_{-\infty}^{x}f(x)=\int_{-\infty}^{x}\frac{1}{\sigma\sqrt{2\pi}}e^{\frac{-(x-\mu)^2}{2\sigma^2}}$  
So, $P(X\leq137.0)=\int_{-\infty}^{137}\frac{1}{10.2\sqrt{2\pi}}e^{\frac{-(137-128)^2}{2\times10.2^2}}=$ `r round(pnorm(137, 128, 10.2), 2)`    
Therefore, 
$P(X>137.0)=1-P(X\leq137.0)=$ **`r round(1-pnorm(137, 128, 10.2), 2)`**

(b)
Let $\bar X$ denote the sampling distribution (n=50). The underlying population distribution is normal, so $\bar X$~$N(\mu, \frac{\sigma^2}{n})$ = $\bar X$~$N(128.0, \frac{10.2}{\sqrt50})$  

The probability of sample mean of blood pressure of 50 males between 20 and 29 yaers old to be less than 125.0 is $P(\bar X<125.0)=$ **`r round(pnorm(125, 128, (10.2)/sqrt(50)), 2)`**

(c)
When sample size is 40, the sampling distribution $\bar X$~$N(128.0, \frac{10.2}{\sqrt40})$   
The 90th percentile of this distribution is **`r round(qnorm(0.9, 128, (10.2)/sqrt(40)), 2)`**

# Problem 4
Let $\bar X$ denote the sample distribution (n=40): $\frac{\bar X-\mu}{s/\sqrt n}$~$t_{(n-1)}$  

(a) The population standard deviation is unknown, so 95% confidence interval for the population mean pulse rate of young females with fibromyalgia is  

$\bar X-t_{n-1, 1-\alpha/2}\frac{s}{\sqrt n}\leq \mu\leq\bar X+t_{n-1, 1-\alpha/2}\frac{s}{\sqrt n}$  
$t_{39,.975}=$ `r round(qt(.975, 39), 2)`  

Therefore, 95% confidence interval is  
$[80-$ `r round(qt(.975, 39), 2)` $(10)/\sqrt 40, 80+$ `r round(qt(.975, 39), 2)` $(10)/\sqrt 40]=$ **(`r round(80-qt(.975, 39)*10/sqrt(40), 2)`, `r round(80+qt(.975, 39)*10/sqrt(40), 2)`)**

(b)
When we repeat drawing samples from the same population, among 95% confidence intervals calculated from each sample, 95% of them has the population mean $\mu$. And (`r round(80-qt(.975, 39)*10/sqrt(40), 2)`, `r round(80+qt(.975, 39)*10/sqrt(40), 2)`) is one of the ranges that was calculated from this particular experiment.

(c)
Let $\mu_0=70$     
$H_0:$ the mean pulse = 70, $H_1:$ the mean pulse $\neq$ 70 ($\alpha=0.01$)  

The test statistic is $t=\frac{\bar X-\mu_o}{s/\sqrt n}=\frac{80-70}{10/\sqrt40}=$ `r round(10/(10/sqrt(40)), 2)`

$t_{n-1, 1-\alpha/2}=$ `r round(qt(.995, 39), 2)`   
$|t|>t_{n-1, 1-\alpha/2}$ so we reject the null hypothesis at a significance level of .01.

Thus, the mean pulse of young women suffering from fibromyalgia does differ significantly from 70.  







